A Generic and Efficient E-field Parallel Imaging Correlator for Next-Generation Radio Telescopes

### 2015-10-24 imaging_demo/ms.tex
fixing some typos

% Abstract of the paper
\begin{abstract}
Modern radio telescopes are marching towards densely packed array layouts 
consisting of a large number of antennas ($N$). Since the cost of traditional 
correlators scales as $\mathcal{O}(N^2)$, there will be a steep cost for 
realizing the full imaging potential of these powerful instruments. Through 
our generic and efficient E-field Parallel Imaging Correlator (EPIC), we present 
the first software demonstration of a generalized direct imaging algorithm 
known by the name Modular Optimal Frequency Fourier (MOFF) imager. It takes 
advantage of the multiplication-convolution theorem of Fourier transform. Not 
only does it bring down the cost to $\mathcal{O}(N\log N)$ but can also image 
from irregularly arranged antenna arrays. EPIC is highly modular and 
parallelizable. It is implemented in object oriented Python and is publicly 
available. We have verified the images produced to be mathematically identical 
to those produced using traditional techniques. We have also validated our 
implementation on data observed with the Long Wavelength Array. Antenna systems
with a dense filling factor consisting of a large number of antennas such as 
LWA, SKA, and HERA will gain a significant advantage by deploying EPIC. 
Inherent availability of calibrated time-domain images on timescales roughly 
equal to the writeout timescale of the digitizer will make it a prime 
candidate for transient searches of primarily Fast Radio Bursts (FRB) as well 
as slower planteary and exoplantenary phenomena. 
\end{abstract}

@@

Here, we discuss the components of these architectures in detail.

@@

The squaring operation under MOFF imaging in the image plane introduces antenna 
auto-correlations around the zero spacing in the $uv$-plane which are absent in 
traditional visibility-based imaging. In order for a robust comparison, these 
auto-correlations have to be removed from both the $uv$ and image planes.

@@

The synthesized beam rms is proportional to the antenna power pattern as 
expected from a point spread function uncorrected for the antenna power 
pattern. The amplitude of the rms of difference of the synthesized beams is 
also modulated by the antenna power pattern. The rms of the difference is 
definitely lesser than the rms of the synthesized beam in the central regions 
up to $(l^2+m^2)^{1/2}\lesssim 0.3$. This implies that the beams are well 
matched in the central regions. In the outer regions, their mismatch is 
comparable to the rms of synthesized beams. This indicates the two 
synthesized beams are not completely randomly different from each other in
which case the rms of the difference would have been $\approx \sqrt{2}$
higher than the rms of the each of the synthesized beams. This indicates 
that while differences exist, large fractions of them are still well matched
to each other.Thus the rounding errors in gridding do not affect the 
statistics of the images or the synthesized beams.

@@

Fig.~\ref{fig:parameter-space-computations-instruments} shows the number of 
complex operations per frequency channel per integration timescale. Telescopes 
that fall to the left of this line indicate MOFF imaging is computationally
more efficient than visibility based imaging. All HERA layouts except possibly
HERA-19 are in a parameter space where MOFF imaging holds the advantage. The 
solid line showing future trajectory of HERA like systems will be clearly 
favoured by MOFF imaging. The gray shaded area is for a projected LWA 
expansion and is also predominantly in the region favouring MOFF imaging. 
It is bounded by the LWA1 and LWA-OV on the left and right 
respectively. The current (see Table~\ref{tab:antenna-layouts} and
an expanded layout with a four-fold increase in number of elements over a
two-fold increase in $b_\textrm{max}$ provide the bounds at the bottom and top 
respectively. Current instruments such as MWA and LOFAR lie in parameter space
favouring visibility based imaging.

>>>>

% Abstract of the paper
\begin{abstract}
Modern radio telescopes are marching towards densely packed array layouts 
consisting of a large number of antennas ($N$). Since the cost of traditional 
correlators scales as $\mathcal{O}(N^2)$, there will be a steep cost for 
realizing the full imaging potential of these powerful instruments. Through 
our generic and efficient E-field Parallel Imaging Correlator (EPIC), we present 
the first software demonstration of a generalized direct imaging algorithm 
known as the Modular Optimal Frequency Fourier (MOFF) imager. It takes 
advantage of the multiplication-convolution theorem of Fourier transforms. Not 
only does it bring down the cost to $\mathcal{O}(N\log N)$ but can also image 
from irregularly arranged heterogeneous antenna arrays. EPIC is highly modular and 
parallelizable. It is implemented in object oriented Python and is publicly 
available. We have verified the images produced to be mathematically identical 
to those produced using traditional techniques. We have also validated our 
implementation on data observed with the Long Wavelength Array. Antenna systems
with a dense filling factor consisting of a large number of antennas such as 
LWA, SKA, and HERA will gain a significant advantage by deploying EPIC. 
Inherent availability of calibrated time-domain images on timescales roughly 
equal to the writeout timescale of the digitizer will make it a prime 
candidate for transient searches of primarily Fast Radio Bursts (FRB) as well 
as slower planteary and exoplantenary phenomena. 
\end{abstract}

@@

Here we discuss the components of these architectures in detail.

@@

The squaring operation under MOFF imaging in the image plane introduces antenna 
auto-correlations around the zero spacing in the $uv$-plane which are absent in 
traditional visibility-based imaging. In order to achieve a robust comparison, these 
auto-correlations have to be removed from both the $uv$ and image planes.

@@

The synthesized beam rms is proportional to the antenna power pattern as 
expected from a point spread function uncorrected for the antenna power 
pattern. The rms of differenced synthesized beams is 
also modulated by the antenna power pattern. The rms of the difference is 
definitely lesser than the rms of the synthesized beam in the central regions 
up to $(l^2+m^2)^{1/2}\lesssim 0.3$. This implies that the beams are well 
matched in the central regions. In the outer regions, their mismatch is 
comparable to the rms of synthesized beams. This indicates the two 
synthesized beams are not completely randomly different from each other in
which case the rms of the difference would have been $\approx \sqrt{2}$
higher than the rms of the each of the synthesized beams. This indicates 
that while differences exist, large fractions of them are still well matched
to each other.Thus the rounding errors in gridding do not affect the 
statistics of the images or the synthesized beams.

@@

Fig.~\ref{fig:parameter-space-computations-instruments} shows the number of 
complex operations per frequency channel per integration timescale. Telescopes 
that fall to the left of this line indicate MOFF imaging is computationally
more efficient than visibility based imaging. All HERA layouts except possibly
HERA-19 are in a parameter space where MOFF imaging holds the advantage. The 
solid line showing future trajectory of HERA like systems will be clearly 
favoured by MOFF imaging. The gray shaded area is for a projected LWA 
expansion and is also predominantly in the region favouring MOFF imaging. 
It is bounded by the LWA1 and LWA-OV on the left and right 
respectively. The current (see Table~\ref{tab:antenna-layouts}) and
an expanded layout with a four-fold increase in number of elements over a
two-fold increase in $b_\textrm{max}$ provide the bounds at the bottom and top 
respectively. Current instruments such as MWA and LOFAR lie in parameter space
favouring visibility based imaging.
